
---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Applying Urinary Biomarkers  8-isoprostane to Understand the Health Effects of NO2 Exposure"
subtitle: "https://github.com/Yang190809/data_repository_yang"
author: "Yang Wang"
fontsize: 12pt
mainfont: Times New Roman

---
**Abstract**

Mounting evidence shows that exposure to NO2 generates oxidative stress. Oxidative stress can cause lipid damage, which plays an important role in various respiratory and cardiovascular diseases.  Urinary 8-isoprostane is a product of lipid peroxidation which can reflect systemic lipid damage. I built linear mix models to analyze the relationship between the level of urinary 8-isoprostane with the level of NO2 exposure. I found that short term NO2 exposure was associated with lipid peroxidation, reflected by increased concentrations of urinary 8-isoprostane associated with increasing exposure. One IQR (7.41 µg/m3) incremental change of 12-h NO2 exposure was associated with an increase of 8-isoprostane level by 19.45% (95% Cl: 14.37%, 24.55%, p-value <0.001). One IQR (8.64 µg/m3) incremental change of 24-h NO2 exposure was associated an increase in 8-isoprostane level by 27.69% (95% Cl: 20.50%, 34.95% ,  p-value <0.001). One IQR (4.47 µg/m3) incremental change of 1-week NO2 exposure was associated with an increase in 8-isoprostane level by 25.15% (95% Cl: 14.14%,36.31%, p-value <0.05). One IQR (2.74 µg/m3) incremental change of 2-week NO2 exposure was associated with an increase in 8-isoprostane level by 15.68% (95% Cl: 8.87%,22.65%, p-value <0.05). 

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 


\newpage
# Rationale and Research Questions
Exposure to NO2 associated with cardiovascular disease, lung function impairment and asthma (Mölter, A., et. al.,2013, Collart, P., et. al.,2018, Takenoue, Y., et. al., 2012 ). Mounting evidence shows that exposure to NO2 generates oxidative stress (Hashemzadeh, B., et. al.,2019). Oxidative stress can cause lipid damage (Black, C. N., et. al., 2017), which plays an important role in various respiratory and cardiovascular diseases. (Zanolin, M. E., et. a.,2015, Lakshmi, S. V., et. al., 2009). I applied 8-isoprostane to investigate the health effects of short term NO2 exposure.


8-isoprostane is the product of the oxidized cell membrane after being attacked by reactive oxygen species such as peroxides. Therefore, 8-isoprostane can reflect lipid damage (Danielsen et al., 2009). 8-isoprostane in urine does not have a diurnal variation and has been proven to be a stable biomarker showing systemic lipid damage. It has been applied in studying diseases such as type 2 diabetes mellitus, obesity, coronary heart disease, asthma, and acute respiratory distress syndrome (Nuernberg, A. M., et al., 2008). Nevertheless, the use of this biomarker in the research of air pollution and its health effects is scarce. One study in Iran discovered a significant positive relationship between short-term NO2 exposure and 8-isoprostane in exhaled breath condensate (EBC) in healthy children aged 12-13 years old (Hashemzadeh, B., et. al., 2019). One study in New York City found that the increases in 1- to 5-day averages of nitrogen dioxide were significantly associated with increases in 8-isoprostane in EBC among 18-year old healthy and asthma affected individuals (Patel, M. M., et. al., 2013).
My research question is:  If the NO2 exposure is positively associated with urinary 8-isoprostane?
Among periods of 12-hour,24-hour,1-week and 2-weeks, which period has the most significant association with urinary 8-isoprostane?


# Dataset Information

##Data Collection
All the data are obtained from Jim Zhang's lab. The urine samples were collected from participants of a previously described study conducted in Changsha, China by Dr. Zhang and other researchers ( Day, D. B., et. al.,2018). The study recruited 89 healthy individuals (age>18years old) who were living and working at the Broad Company Campus and divided them into 2 groups, Group A (n = 36) and Group B (n = 53). The study periods include pre-intervention, intervention, and post-intervention, lasting 5 weeks. Urine samples were collected from each participant once during the pre-intervention, twice during the intervention, and once during the post-intervention. The study measured hourly indoor and outdoor concentrations of PM2.5, O3, NO2, and SO2, surveyed each participant’s time-activity, and calculated the participants’ exposure for each pollutant over 12h, 24h, 1 week and 2 week periods, which were counted backward from the visit points.

The file 8-iso is the data set of urinary biomarker 8-isoprostane.The file urine_info is the data set that has information about the subjects who provided the urine samples. The urine list is a data set with lists of information of sample ID, subject ID and visits. The meaning of the columns as well as units and class of the data in each foler is listed below. Column names without descriptors are irrelevant to this study.

##Urine List
The urine list is a data set with lists of information of sample ID, subject ID and visits. 

Column Name	| Meaning	                            |units |	  class of the data
------------|-------------------------------------|----- |---------------------
 Sample ID	|   the identity number of the samples	  |NA	 |      integer
Subject ID |	 the identity number of the subjects	|NA	    |   integer
 visit	  |   the number of the 4 visits (1,2,3 or 4)	|NA	  |     integer
kable(Table1,caption="Urine list raw data set") 

##8-isoprostane Data Set
The 8-isoprostane data set has information about 8-isoprostane concentration tested in the mass spectrum and the concentration of 8-isoprostane in original urine samples. The limit of detection for 8-isoprostane was 0.016ng/ml. Any value which is below 0.016 in the column of calculated Conc in 8-is data set should be excluded as an error.

Column Name	|    Meaning                                   |    Units	  |    Class of the data
------------|----------------------------------------------|------------|---------------------
  Sample ID |      the identity number of the samples	|         NA	   |     integer
 Calculated Conc|	the concentration tested by the machine   |   ng/ml	 |   numeric
Sample Conc	 |   the concentration in the original urine|	     ng/ml	  |  numeric

kable(Table2,caption="8-isoprostane data set ") 

##Urine Info Data Set
The urine info is the data set that has information about subjects characteristics and average pollutant exposure (NO2, SO2, O3, and PM2.5)over the periods of 12 hours, 24 hours, one week, and two weeks. It also includes average temperature and humidity over the periods of 12 hours, 24 hours, one week, and two weeks. 


Column name|	Meaning                             |	Units|	Class of the data
---------- |--------------------------------------|----- |---------------------
ample_ID  |	the identity number of the samples  |	NA |	integer            
SubjectID  |	the identity number of the subjects| NA	|integer    
COLD	|cold ( represent respiratory infection)|	NA|	category
MNST	|menstration during visit|	NA|	category
last.ate|	the hours to the last meal|	hours	|integer
wkday.start|	the day that the subject start their work	|NA	|category
dt_smoke|	second-hand smoke exposure in hours|	hours	|numeric
USG	urine |specific gravity	|g/ml	|numeric
o3exp.12h|	the exposure of ozone in 12h|	ug/m3	|numeric
pmexp.12h	|the exposure of PM2.5 in 12h|	ug/m3|	numeric
no2exp.12h|	the exposure of NO2 in 12h|	ug/m3	|numeric
so2exp.12h|	the exposure of SO2 in 12h	|ug/m3|	numeric
Temp.12h|	temperature in 12h|	ug/m3	|numeric
RHx.12h|	humidity in 12h	|ug/m3|	numeric

kable(Table3,caption="Urine info	data set") 


## Data Wrangling
My data wrangling started with loading the data set of 8-isoprostane and changing the column names. Then I merged this data set with the urine list to match the sample ID with subjects ID and visit. Then this merged data was merged with the urine info data set using the subject ID and visit as the matching keys. After the second merging, the rows with NAs were removed. Then I calculated the normalized 8-isoprostane using the specific urine gravity. This normalization can adjust the dilution in the urine. I obtained my final data by selecting columns that are relevant to my research question. Finally, the processed data was saved into a processed data folder.

\newpage

**1. Data wrangling**
```{r 4, warning=FALSE, message=FALSE}
# Set working directory
getwd()
# Load packages
library(tidyverse)
library(lmerTest)
library(MuMIn)
library(car)
library(tidyverse)
library(cowplot)
library("stargazer")
library(knitr)
```


```{r 6, warning=FALSE, message=FALSE}
# Set ggplot theme
mytheme <- theme_classic(base_size = 25) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

# Load dataset 1
is <- read.csv("raw_data/8iso.csv")
is <- select(is, Sample.ID, Calculated.Conc, Sample.Conc)
colnames(is)[colnames(is) == "Calculated.Conc"] <- "is.origin"
colnames(is)[colnames(is) == "Sample.Conc"] <- "is.conc"

# Load dataset 2
list <- read.csv("raw_data/urine_list.csv")

#merge
merge1 <- merge(x=is,y=list,by="Sample.ID",all=TRUE)

urine.info <- read.csv("raw_data/urine_info.csv")
merge3 <- merge(x=merge1,y=urine.info,by=c("Subject.ID","visit"),all=TRUE)
dim(merge3)
full.data <- na.omit(merge3)
dim(full.data)
mean.cre<- mean(full.data$USG)
mean.cre

#calculate normalized biomarker concentration
full.data$is.ad <- (full.data$is.conc/(1-full.data$USG))*(1-mean.cre)

#select useful colums
final.dat<- select(full.data, Sample.ID, Subject.ID, is.ad, is.conc, is.origin, gender,	COLD,	MNST,	last.ate,	wkday.start,go.home, dt_smoke,	o3exp.12h,	pmexp.12h,no2exp.12h,	so2exp.12h,	Temp.12h,	RHx.12h,	o3exp.24h,	pmexp.24h,	no2exp.24h,	so2exp.24h,	Temp.24h,	RHx.24h,	o3exp.1w,	pmexp.1w,	no2exp.1w,	so2exp.1w,	Temp.1w,	RHx.1w,	o3exp.2w,	pmexp.2w,	no2exp.2w,	so2exp.2w,	Temp.2w,	RHx.2w )

#save processed data set
write.csv(final.dat, row.names = FALSE, file = "processed_data/final.dat.csv")
```

##Explore the data
```{r 7,warning=FALSE, message=FALSE}

#load processed data set
dat <- read.csv("processed_data/final.dat.csv")
#remove outliers
summary(dat$is.origin)
dat1<- dat[-c(4,7,89,147,167),]

#test normality
shapiro.test(dat1$is.ad)
shapiro.test(log(dat1$is.ad))
#explore the data
```

```{r 8, include=FALSE}
is.sum <- dat1 %>%
  summarise(mean = mean(is.ad, na.rm = TRUE),
            minimum = min(is.ad, na.rm = TRUE),
            maximum = max(is.ad, na.rm = TRUE),
            Standard.dev= sd(is.ad, na.rm = TRUE))

log.is.sum<- dat1 %>%
  summarise(mean = mean(log(is.ad), na.rm = TRUE),
            minimum = min(log(is.ad), na.rm = TRUE),
            maximum = max(log(is.ad), na.rm = TRUE),
            Standard.dev= sd(log(is.ad), na.rm = TRUE))

no2.sum1<-summary(dat1$no2exp.12h) 
no2.sum2<-summary(dat1$no2exp.24h)
no2.sum3<-summary(dat1$no2exp.1w)
no2.sum4<-summary(dat1$no2exp.2w)
```

```{r 9, echo=FALSE}
sub.dat<- dat1 %>%
  select(no2exp.12h,no2exp.24h,no2exp.1w,no2exp.2w)
table4<-stargazer(sub.dat ,type = "text", 
            title = "Summary of Statistics",
            covariate.labels = c("12-hour NO2 exposure","24-hour NO2 exposure","1-week NO2 exposure","2-week NO2 exposure"),
            style = "qje",
            out="sum.no2")
kable(table4 ,caption="Summary of NO2 exposure") 
```



```{r 11,warning=FALSE, message=FALSE}
#calculate IQR for each period
no2.12 <-no2.sum1[5]-no2.sum1[2]
no2.12
no2.24 <-no2.sum2[5]-no2.sum2[2]
no2.24
no2.1w <-no2.sum3[5]-no2.sum3[2]
no2.1w
no2.2w <-no2.sum4[5]-no2.sum4[2]
no2.2w
```

**3.Build models**

3.1 Build models for 12-hour NO2 exposure
```{r 12,warning=FALSE, message=FALSE}
mo1<- lmer(log(is.ad) ~no2exp.12h+o3exp.12h+ pmexp.12h+so2exp.12h+Temp.12h+RHx.12h + last.ate +wkday.start+go.home+COLD+MNST+dt_smoke +(1|Subject.ID),data=dat1)
#find the best model with lowest AIC
step(mo1)
#best model for 12-hour NO2 exposure
mo1.1<- lmer(log(is.ad) ~ no2exp.12h +o3exp.12h + (1 | Subject.ID),data=dat1)
summary(mo1.1)
#check F value
anova(mo1.1)
#check collinearirty
vif(mo1.1)
#get R^2
r.squaredGLMM (mo1.1)
#identify outliers
res1.1 <- resid(mo1, type = "pearson")
dat[which(abs(res1.1) > 2.5),]
#calculate IQR change
(exp(0.025912)-1)*no2.12
(exp(0.025912 +0.006693)-1)*no2.12
(exp(0.025912 -0.006693)-1)*no2.12

```


3.2 Build models for 24-hour NO2 exposure
```{r 13,warning=FALSE, message=FALSE}
mo2<- lmer(log(is.ad) ~ no2exp.24h+o3exp.24h+ pmexp.24h+so2exp.24h+Temp.24h+RHx.24h + last.ate +wkday.start+go.home+COLD+MNST+dt_smoke +(1|Subject.ID),data=dat1)
step(mo2)
mo2.1<- lmer(log(is.ad) ~ no2exp.24h + o3exp.24h + (1 | Subject.ID),data=dat1)
summary(mo2.1)
anova(mo2.1)
vif(mo2.1)
r.squaredGLMM (mo2.1)
res2.1 <- resid(mo2.1, type = "pearson")
dat[which(abs(res2.1) > 2.5),]
#calculate IQR change
(exp(0.031555)-1)*no2.24
(exp(0.031555 +0.008104)-1)*no2.24
(exp(0.031555 -0.008104)-1)*no2.24
```

3.3 Build models for 1-week NO2 exposure
```{r 14 ,warning=FALSE, message=FALSE}
mo3<- lmer(log(is.ad) ~ no2exp.1w+o3exp.1w+ pmexp.1w+so2exp.1w+Temp.1w+RHx.1w + last.ate +wkday.start+go.home+COLD+MNST+dt_smoke +(1|Subject.ID),data=dat1)
step(mo3)
mo3.1<- lmer(log(is.ad) ~ no2exp.1w + so2exp.1w + (1 | Subject.ID),data=dat1)
summary(mo3.1)
anova(mo3.1)
vif(mo3.1)
r.squaredGLMM (mo3.1)
#calculate IQR change
(exp( 0.02877)-1)*no2.1w
(exp( 0.02877 +0.01250)-1)*no2.1w
(exp( 0.02877 -0.01250)-1)*no2.1w

```

3.4 Build models for 2-week NO2 exposure
```{r 15,warning=FALSE, message=FALSE}

mo4<- lmer(log(is.ad) ~ no2exp.2w+o3exp.2w+ pmexp.2w+so2exp.2w+Temp.2w+RHx.2w + last.ate +wkday.start+go.home+COLD+MNST+dt_smoke +(1|Subject.ID),data=dat1)
summary(mo4)
step(mo4)
mo4.1<- lmer(log(is.ad) ~ no2exp.2w + so2exp.2w + (1 | Subject.ID),data=dat1)
summary(mo4.1)
anova(mo4.1)
vif(mo4.1)
r.squaredGLMM (mo4.1)
#calculate IQR change
(exp(0.05572)-1)*no2.2w
(exp(0.05572 +0.02382)-1)*no2.2w
(exp(0.05572 -0.02382)-1)*no2.2w

```

# Exploratory Analysis 
The Shapiro tests showed that concentrations of urinary 8-isoprostane are not normalized (p-value < 2.2e-16). The Shapiro tests showed that concentrations of log-scale urinary 8-isoprostane are still not normalized (p-value < 2.2e-16). To build linear models, I need to use the more normalized log-transformed concentrations of urinary 8-isoprostane. The distribution of data in log(8-isoprostane) (Figure 2)looks more normalized than the distribution of data of 8-isoprostane in (Figure 1).

```{r 16,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Plot  of  8-isoprostane"}
is.plot1 <- boxplot(dat1$is.ad,colors="darkblue",ylab="8-isoprostane ng/ml")
```

```{r 17,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Plot  of  log(8-isoprostane)"}
is.plot2 <- boxplot(log(dat1$is.ad),color="darkblue",ylab="log(8-isoprostane ng/ml)")
```


# Analysis
I used linear mixed models with participant-specific intercepts. The inclusion of participants as a variable in mixed-effects models account for the correlation of repeated measurements from the same individuals and precludes the need to control for participant characteristics (e.g., age, gender, BMI, the identity of smoker or non-smoker) that do not change across the four longitudinal measurements.

I used NO2 exposure as predicting variables and log-transformed urinary biomarkers as dependent variables. For each biomarker, I built 4 models to the exposure over the periods of 12h, 24h, one week, and two weeks. I used a backward stepwise model selection method to select the co-variables for each model. The co-variables that I tested in the models were last meal, the start of the workday, respiratory infection status, menstruation, and the time of second-hand exposure. I also tested the SO2 exposure, ozone exposure and PM2.5 exposure, the average temperature and humidity during a corresponding period. 


With the stepAIC function, the best models were chosen with the lowest AIC. The vif tests show that there is no model with excess intercorrelation in the predicting variables. Figure 3 shows that all four diagnostic plots have the shape of a triangle. The higher the fitted value is, the higher the residue is. This residue is a result of skewed distribution of log scaled of the predicting variable .In general, the residue distribution is not bad. No value is bigger than an absolute value of 2.5.
```{r 18,warning=FALSE, message=FALSE,include=FALSE}
p1<-plot(mo1.1)
p2<-plot(mo2.1)
p3<-plot(mo3.1)
p4<-plot(mo1.1)
```

```{r 19,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Diagnostic plots of four final models"}
plot_grid(p1,p2,p3,p4, nrow = 2, align = 'h')
```
Both 12-hour, 24-hour, 1-week and 2-week NO2 exposure showed significant correlations with the level of urinary 8-isoprostane, especially the 2-week NO2 exposure. 

## Result 1
```{r 20 ,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Relationship between log(8-isoprostane) and 12-hour NO2 exposure"}
is.12h <- ggplot(dat1, aes(y =log(is.ad) , x =no2exp.12h,color=dat1$gender)) +
   geom_point(shape=1, size = 2) +
  geom_smooth(method="lm",se=FALSE,color="black")+
  labs(x = "NO2 exposure in 12 hours", y = "log(8-isoprostane ng/ml)") +
    theme(legend.position="top")+
  ylim(0,5)
print(is.12h)
```

There is a significant positive relationship between 12-hour NO2 exposure and log(8-isoprostane)(Figure 7).One IQR (7.41 µg/m3) incremental change of 12-h NO2 exposure was associated with an increase of 8-isoprostane level by 19.45% ( 95% Cl: 14.37%, 24.55%, F value=14.99, p-value <0.001). The Adjusted R-squared =0.4121, which is the fraction of total variance explained by the model

## Result 2
```{r 21,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Relationship between log(8-isoprostane) and 24-hour NO2 exposure"}
is.24h <- ggplot(dat1, aes(y =log(is.ad) , x =no2exp.24h,color=dat1$gender)) +
   geom_point(shape=1, size = 2) +
  geom_smooth(method="lm",se=FALSE,color="black")+
  labs(x = "NO2 exposure in 24 hours", y = "log(8-isoprostane ng/ml)") +
    theme(legend.position="top")+
  ylim(0,5)
print(is.24h)
```


There is a signifcant positive relationship between 24-hour NO2 exposure and log(8-isoprostane)(Figure 8).One IQR (8.64 µg/m3) incremental change of 24-h NO2 exposure was associated an increase in 8-isoprostane level by 27.69% ( 95% Cl: 20.50%, 34.95% , F value=15.16,  p-value <0.001 ). The Adjusted R-squared = 0.4160, which is the fraction of total variance explained by the model.


## Result 3
```{r 22,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Relationship between log(8-isoprostane) and 1-week NO2 exposure"}
is.1w <- ggplot(dat1, aes(y =log(is.ad) , x =no2exp.1w,color=dat1$gender)) +

   geom_point(shape=1, size = 2) +
  geom_smooth(method="lm",se=FALSE,color="black")+
  labs(x = "NO2 exposure in one week", y = "log(8-isoprostane ng/ml)") +
    theme(legend.position="top")+
  ylim(0,5)
print(is.1w)
```


There is a significant positive relationship between 1-week NO2 exposure and log(8-isoprostane)(Figure 9).One IQR (4.47 µg/m3) incremental change of 1-week NO2 exposure was associated with an increase in 8-isoprostane level by 25.15% ( 95% Cl: 14.14%,36.31%, F value=5.30, p-value <0.05 ). The Adjusted R-squared = 0.4014, which is the fraction of total variance explained by the model.


## Result 4
```{r 23,warning=FALSE, message=FALSE,echo=FALSE,fig.cap="Relationship between log(8-isoprostane) and 2-week NO2 exposure"}
is.2w <- ggplot(dat1, aes(y =log(is.ad) , x =no2exp.2w,color=dat1$gender)) +
   geom_point(shape=1, size = 2) +
  geom_smooth(method="lm",se=FALSE,color="black")+
  labs(x = "NO2 exposure in 2 weeks", y = "log(8-isoprostane ng/ml)") +
    theme(legend.position="top")+
  ylim(0,5)
print(is.2w)

```

There is a significant positive relationship between 2-week NO2 exposure and log(8-isoprostane)(Figure 10).One IQR (2.74 µg/m3) incremental change of 2-week NO2 exposure was associated with an increase in 8-isoprostane level by 15.68% ( 95% Cl: 8.87%,22.65%, F value=5.47, p-value <0.05 ). The Adjusted R-squared = 0.4021, which is the fraction of total variance explained by the model.

# Summary and Conclusions
Short term NO2 exposure was associated with urinary 8-isoprostane. This finding meet my hypothesis. Urinary 8-isoprostane indicates the lipid peroxidation from the whole body, reflecting the systemic oxidative stress. It indicates short term NO2 exposure can cause signifiant higher systemic oxidative stress. 

\newpage

# References
Mölter, A., Agius, R. M., de Vocht, F., Lindley, S., Gerrard, W., Lowe, L., ... & Simpson, A. (2013). Long-term exposure to PM10 and NO2 in association with lung volume and airway resistance in the MAAS birth cohort. Environmental health perspectives, 121(10), 1232-1238.
Collart, P., Dubourg, D., Levêque, A., Sierra, N. B., & Coppieters, Y. (2018). Short-term effects of nitrogen dioxide on hospital admissions for cardiovascular disease in Wallonia, Belgium. International journal of cardiology, 255, 231-236.
Takenoue, Y., Kaneko, T., Miyamae, T., Mori, M., & Yokota, S. (2012). Influence of outdoor NO2 exposure on asthma in childhood: Meta‐analysis. Pediatrics International, 54(6), 762-769.
Hashemzadeh, B., Idani, E., Goudarzi, G., Ankali, K. A., Sakhvidi, M. J. Z., Babaei, A. A., ... & Neisi, A. (2019). Effects of PM2. 5 and NO2 on the 8-isoprostane and lung function indices of FVC and FEV1 in students of Ahvaz city, Iran. Saudi journal of biological sciences, 26(3), 473-480.
Black, C. N., Bot, M., Révész, D., Scheffer, P. G., & Penninx, B. (2017). The association between three major physiological stress systems and oxidative DNA and lipid damage. Psychoneuroendocrinology, 80, 56-66.
Lakshmi, S. V., Padmaja, G., Kuppusamy, P., & Kutala, V. K. (2009). Oxidative stress in cardiovascular disease.
Zanolin, M. E., Chamitava, L., Degan, P., Pasini, A., Fratta-Pasini, A., Nicolis, M., ... & De Marco, R. (2015). Biomarkers of oxidative stress in chronic respiratory diseases.
Danielsen, P. H., Loft, S., Kocbach, A., Schwarze, P. E., & Møller, P. (2009). Oxidative damage to DNA and repair induced by Norwegian wood smoke particles in human A549 and THP-1 cell lines. Mutation Research/Genetic Toxicology and Environmental Mutagenesis, 674(1-2), 116-122.
Nuernberg, A. M., Boyce, P. D., Cavallari, J. M., Fang, S. C., Eisen, E. A., & Christiani, D. C. (2008). Urinary 8-isoprostane and 8-OHdG concentrations in boilermakers with welding exposure. Journal of occupational and environmental medicine, 50(2), 182-189.
Hashemzadeh, B., Idani, E., Goudarzi, G., Ankali, K. A., Sakhvidi, M. J. Z., Babaei, A. A., ... & Neisi, A. (2019). Effects of PM2. 5 and NO2 on the 8-isoprostane and lung function indices of FVC and FEV1 in students of Ahvaz city, Iran. Saudi journal of biological sciences, 26(3), 473-480.
Patel, M. M., Chillrud, S. N., Deepti, K. C., Ross, J. M., & Kinney, P. L. (2013). Traffic-related air pollutants and exhaled markers of airway inflammation and oxidative stress in New York City adolescents. Environmental research, 121, 71-78.
